---
title: "Hydrological Analysis Report"
author: "Samuel Brhane Alemayohu"
date: "2024-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("corrplot")
# install.packages("writexl")

library(writexl)
library(readxl)
library(dplyr)
library(ggplot2)
library(corrplot)
```

```{r echo=FALSE}
gauge_stations <- list(
  Luče = list(longitude = 14.7457, latitude = 46.3543, elevation = 513),
  Solčava = list(longitude = 14.6914, latitude = 46.4203, elevation = 639),
  Laško = list(longitude = 15.2341, latitude = 46.1565, elevation = 221),
  GornjiGrad = list(longitude = 14.8077, latitude = 46.2960, elevation = 430),
  Radegunda = list(longitude = 14.9330, latitude = 46.3661, elevation = 794)
)

temp_station <- list(
  CeljeMedlog = list(longitude = 15.2259, latitude = 46.2366, elevation = 242)
)
```

# [1. DATA PREPARATION]{style="color:blue; font-family:Arial; font-size:28px"}

[In this section, we import the hydrological data and perform initial processing, which includes checking for any missing values and cleaning the dataset. If missing data is found, appropriate methods will be used to address them.]{style="font-family:Arial; font-size:16px;"}

```{r data_import}
# Import dataset from excel file
main_data = read_excel("Savinja-VelikoSirje.xlsx")

# Create dataframe
main_data = as.data.frame(main_data)

# Convert the 'Date' column to Date type
main_data$Date <- as.Date(main_data$Date, format = "%d.%m.%Y")

head(main_data)
```

```{r cleaning_data}
# Structure of dataframe
str(main_data)

# Check for missing values
missing_data <- colSums(is.na(main_data))
missing_data
```

# [1.1. Fill Missing Waterlevel Values]{style="color:#12a23a; font-size:20px"}

```{r waterlevel_imputation}
sum(is.na(main_data$Waterlevel))

# Fit linear regression model using Discharge
model <- lm(Waterlevel ~ Discharge, data = main_data, na.action = na.exclude)
model

# Predict missing Waterlevel
missing_waterlevel <- which(is.na(main_data$Waterlevel))
predicted_values <- predict(model, newdata = main_data[missing_waterlevel, ])
predicted_values

# Fill missing Waterlevel values
main_data$Waterlevel[missing_waterlevel] <- predicted_values

# Check Waterlevel again
sum(is.na(main_data$Waterlevel))
```

# [1.2. Fill Missing Watertemp Values]{style="color:#12a23a; font-size:20px"}

```{r watertemp_imputation}
missing_watertemps = which(is.na(main_data$Watertemp))
missing_watertemps

# Loop missing watertemp and fill 
for (i in missing_watertemps){
  current_year <- main_data$Year[i]
  date_without_year <- format(main_data$Date[i], "%m-%d")
  
  # Find Watertemp and Airtemp for same date in previous years
  previous_data <- main_data %>% 
    filter(format(Date, "%m-%d") == date_without_year & Year < current_year) %>% 
    select(Watertemp, Airtemp)
  
  # Calculate mean of Watertemp and Airtemp
  if(nrow(previous_data) > 0){
    watertemp_mean <- mean(previous_data$Watertemp)
    airtemp_mean <- mean(previous_data$Airtemp)
    
    main_data$Watertemp[i] <- watertemp_mean + (main_data$Airtemp[i] - airtemp_mean)
  }
}

# Check Watertemp
sum(is.na(main_data$Watertemp))
```

# [1.3. Fill Missing Values for P3_Lasko]{style="color:#12a23a; font-size:20px"}

```{r fill_p3_lasko}
missing_lasko <-  which(is.na(main_data$P3_Lasko))

# Fill missing values with average of other stations
for (i in missing_lasko){
  other_prec <- main_data[i, c("P1_Luce", "P2_Solcava", "P4_GornjiGrad", "P5_Radegunda")]
  main_data$P3_Lasko[i] <- mean(as.numeric(other_prec))
}

# Check P3_Lasko
sum(is.na(main_data$P3_Lasko))
```

```{r}
# Check main_data for any missing values
colSums(is.na(main_data))
```

# [2. BASIC ANALYSIS]{style="color:blue; font-family:Arial; font-size:28px"}

[In this section, we perform basic hydrological analysis. This includes calculating descriptive statistics for key variables such as discharge, precipitation, evapotranspiration, and air temperature. We will also plot graphs for these variables over the entire period and for the year 2022. Additionally, we will calculate correlations between precipitation stations, water balance, runoff coefficient, and the ratio between evapotranspiration and precipitation.]{style="font-family:Arial; font-size:16px;"}

# [2.1. Calculate Descriptive Statistics]{style="color:#12a23a; font-size:20px"}

```{r discharge_statistics}
# Calculate descriptive statistics for Discharge
discharge_stats <- data.frame(
  Statistic = c("Mean", "Standard Deviation", "Minimum", "Maximum", "Median"),
  Values = c(
    mean(main_data$Discharge),
    sd(main_data$Discharge),
    min(main_data$Discharge),
    max(main_data$Discharge),
    median(main_data$Discharge))
) 

discharge_stats

# Summary statistics for Discharge
summary(main_data$Discharge)
```
```{r precipitation_statistics}
# Calculate descriptive statistics for Precipitation 
precipitation_stats <- data.frame(
  Statistic = c("Mean", "Standard Deviation", "Minimum", "Maximum", "Median"),
  Values = c(
    mean(c(main_data$P1_Luce, main_data$P2_Solcava, main_data$P3_Lasko, main_data$P4_GornjiGrad, main_data$P5_Radegunda)),
    sd(c(main_data$P1_Luce, main_data$P2_Solcava, main_data$P3_Lasko, main_data$P4_GornjiGrad, main_data$P5_Radegunda)),
    min(c(main_data$P1_Luce, main_data$P2_Solcava, main_data$P3_Lasko, main_data$P4_GornjiGrad, main_data$P5_Radegunda)),
    max(c(main_data$P1_Luce, main_data$P2_Solcava, main_data$P3_Lasko, main_data$P4_GornjiGrad, main_data$P5_Radegunda)),
    median(c(main_data$P1_Luce, main_data$P2_Solcava, main_data$P3_Lasko, main_data$P4_GornjiGrad, main_data$P5_Radegunda))
  )
)

precipitation_stats

# Summary statistics for Precipitation
summary(c(main_data$P1_Luce, main_data$p2_Solcava, main_data$P3_Lasko, main_data$P4_GornjiGrad, main_data$P5_Radegunda))
```
``` {r precipiation_stations}
# P1_Luce
mean(main_data$P1_Luce)
sd(main_data$P1_Luce)
min(main_data$P1_Luce)
max(main_data$P1_Luce)
median(main_data$P1_Luce)

# P2_Solcava
mean(main_data$P2_Solcava)
sd(main_data$P2_Solcava)
min(main_data$P2_Solcava)
max(main_data$P2_Solcava)
median(main_data$P2_Solcava)

# P3_Lasko
mean(main_data$P3_Lasko)
sd(main_data$P3_Lasko)
min(main_data$P3_Lasko)
max(main_data$P3_Lasko)
median(main_data$P3_Lasko)

# P4_GornjiGrad
mean(main_data$P4_GornjiGrad)
sd(main_data$P4_GornjiGrad)
min(main_data$P4_GornjiGrad)
max(main_data$P4_GornjiGrad)
median(main_data$P4_GornjiGrad)

# P5_Radegunda
mean(main_data$P5_Radegunda)
sd(main_data$P5_Radegunda)
min(main_data$P5_Radegunda)
max(main_data$P5_Radegunda)
median(main_data$P5_Radegunda)

```
```{r airtemp_statistics}
# Calculate descriptive statistics for Air Temperature
airtemp_stats <- data.frame(
  Statistic = c("Mean", "Standard Deviation", "Minimum", "Maximum", "Median"),
  Values = c(
    mean(main_data$Airtemp),
    sd(main_data$Airtemp),
    min(main_data$Airtemp),
    max(main_data$Airtemp),
    median(main_data$Airtemp)
  )
)

airtemp_stats

# Summary statistics for Air Temperature
summary(main_data$Airtemp)
```

```{r evapotrans_statistics}
# Calculate descriptive statistics for Evapotranspiration
evapotrans_stats <- data.frame(
  Statistic = c("Mean", "Standard Deviation", "Minimum", "Maximum", "Median"),
  Values = c(
    mean(main_data$Evapotrans),
    sd(main_data$Evapotrans),
    min(main_data$Evapotrans),
    max(main_data$Evapotrans),
    median(main_data$Evapotrans)
  )
)

evapotrans_stats

# Summary statistics for Evapotranspiration
summary(main_data$Evapotrans)
```

# [2.2. Plot Graphs for Key Variables]{style="color:#12a23a; font-size:20px"}

```{r plot_discharge}
# Plot Discharge over the whole period
plot(main_data$Date, main_data$Discharge, type = "l",
     xlab = "Date", ylab = "Discharge (m^3/s)", main = "Discharge Over Whole Period")

# Plot Discharge for year 2022
data_2022 <- main_data[format(main_data$Date, "%Y") == "2022",]
plot(data_2022$Date, data_2022$Discharge, type = "l",
     xlab = "Date", ylab = "Discharge (m^3/s)", main = "Discharge For Year 2022")
```

```{r plot_precipitation}
# Plot Precipitation over whole period for all stations
plot(main_data$Date, main_data$P1_Luce, type = "l", col = "blue",
     xlab = "Date", ylab = "Precipitation (mm)", main = "Precipitation Over Whole Period")

# All lines for other stations
lines(main_data$Date, main_data$P2_Solcava, col = "red")
lines(main_data$Date, main_data$P3_Lasko, col = "green")
lines(main_data$Date, main_data$P4_GornjiGrad, col = "purple")
lines(main_data$Date, main_data$P5_Radegunda, col = "orange")

# Add legend
legend("topright", legend = c("P1_Luce", "P2_Solcava", "P3_Lasko", "P4_GornjiGrad", "P5_Radegunda"),
       col = c("blue", "red", "green", "purple", "orange"), lty = 1)


# Plot Precipitation for year 2022 
plot(data_2022$Date, data_2022$P1_Luce, type = "l", col = "blue",
     xlab = "Date", ylab = "Precipitation (mm)", main = "Precipitation for Year 2022")

# Add lines for other stations
lines(data_2022$Date, data_2022$p2_Solcava, col = "red")
lines(data_2022$Date, data_2022$P3_Lasko, col = "green")
lines(data_2022$Date, data_2022$P4_GornjiGrad, col = "purple")
lines(data_2022$Date, data_2022$P5_Radegunda, col = "orange")

# Add a legend
legend("topright", legend = c("P1_Luce", "P2_Solcava", "P3_Lasko", "P4_GornjiGrad", "P5_Radegunda"),
       col = c("blue", "red", "green", "purple", "orange"), lty = 1)
```
```{r average_precipiation}
# Calculate the average precipitation across all stations
main_data$Avg_Precipitation <- rowMeans(main_data[, c("P1_Luce", "P2_Solcava", "P3_Lasko", "P4_GornjiGrad", "P5_Radegunda")])

# Plot Average Precipitation over the whole period
plot(main_data$Date, main_data$Avg_Precipitation, type = "l",
     xlab = "Date", ylab = "Average Precipitation (mm)", main = "Average Precipitation Over Whole Period")

# Plot Average Precipitation for year 2022
data_2022 <- main_data[format(main_data$Date, "%Y") == "2022",]
plot(data_2022$Date, data_2022$Avg_Precipitation, type = "l",
     xlab = "Date", ylab = "Average Precipitation (mm)", main = "Average Precipitation For Year 2022")
```

```{r plot_evapotranspiration}
# Plot Evapotranspiration over whole period
plot(main_data$Date, main_data$Evapotrans, type = "l", col = "blue",
     xlab = "Date", ylab = "Evapotranspiration (mm)", main = "Evapotranspiration Over Whole Period")

# Plot Evapotranspiration for year 2022
data_2022 <- main_data[format(main_data$Date, "%Y") == "2022",]
plot(data_2022$Date, data_2022$Evapotrans, type = "l", col = "blue",
     xlab = "Date", ylab = "Evapotranspiration (mm)", main = "Evapotranspiration For Year 2022")
```

```{r plot_airtemprature}
# Plot Air Temperature over the whole period
plot(main_data$Date, main_data$Airtemp, type = "l", col = "red",
     xlab = "Date", ylab = "Air Temperature (°C)", main = "Air Temperature Over Whole Period")

# Plot Air Temperature for year 2022
plot(data_2022$Date, data_2022$Airtemp, type = "l", col = "red",
     xlab = "Date", ylab = "Air Temperature (°C)", main = "Air Temperature For Year 2022")
```


# [2.3. Calculate Correlation Between Precipitation Stations]{style="color:#12a23a; font-size:20px"}
```{r plot_correlation}
# Precipitation columns
precipitation_data <- main_data[, c("P1_Luce", "P2_Solcava", "P3_Lasko", "P4_GornjiGrad", "P5_Radegunda")]

# Calculate correlation matrix
cor_matrix <- cor(precipitation_data)

cor_matrix

# Plot correlogram
corrplot(cor_matrix, method = "shade", type = "full", 
         tl.col = "orange", tl.srt = 45,
         title = "Correlogram of Precipitation Stations", 
          mar = c(0, 0, 2, 0)
)
```


# [2.4. Calculate Water Balance]{style="color:#12a23a; font-size:20px"}

```{r water_balance}
# Catchment area in m²
catchment_area <- 1847000000

# Calculate Runoff
main_data$Runoff <- (main_data$Discharge * 3600 * 24 * 1000) / catchment_area

# Calculate the water balance
main_data$Water_Balance <- main_data$Avg_Precipitation - main_data$Evapotrans - main_data$Runoff

# Save main_data to an Excel file
# write_xlsx(main_data, "main_data_output.xlsx")
```

# [2.5. Calculate Runoff Coefficient ]{style="color:#12a23a; font-size:20px"}

```{r runoff_coeffiecient}
# Group data by Year and calculate total runoff and total precipitation
annual_runoff <- main_data %>%
  group_by(Year) %>%
  summarise(
    Total_Runoff = sum(Runoff),             
    Total_Precipitation = sum(Avg_Precipitation)
  )

# Calculate the runoff coefficient for each year
annual_runoff$Runoff_Coefficient <- annual_runoff$Total_Runoff / annual_runoff$Total_Precipitation

annual_runoff

# Plot the runoff coefficient over the years
plot(as.numeric(annual_runoff$Year), annual_runoff$Runoff_Coefficient, type = "b", col = "blue",
     xlab = "Year", ylab = "Runoff Coefficient", main = "Runoff Coefficient Values")

```


# [2.6. Calculate Ratio Between Evapotranspiration and Precipitation  ]{style="color:#12a23a; font-size:20px"}

```{r evapotrans_prec_ratio}
# Group data by Year and calculate total evapotranspiration and total precipitation
evapotrans_prec <- main_data %>% 
  group_by(Year) %>% 
  summarise(
    Avg_Evapotrans = sum(Evapotrans),
    Avg_Precipitation = sum(Avg_Precipitation))

# Calculate evapotranspiration to precipitation ratio
evapotrans_prec$Ratio <- evapotrans_prec$Avg_Evapotrans / evapotrans_prec$Avg_Precipitation

evapotrans_prec

# Plot evapotranpitation to precipitation ratio
plot(evapotrans_prec$Year, evapotrans_prec$Ratio, type = "b", col = "orange",
     xlab = "Year", ylab = "Ratio", main = "Evapotranspiration to Precipitation Ratio")
```

